A multi-scalar story of the diffusion of a new technology: the web

### 2024-02-08 paper/submission1/tranos2023.qmd
bua

<!-- motivation -->
The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffused over space and time and use this knowledge to make predictions
regarding the diffusion of related *future* technologies.
As per @leibowicz2016representing, historical studies agree on the fact that 
"technologies diffuse at different times, at different rates, and to
different extents in different places, and can be significantly
influenced by policies [@victor1993]". @meade2021modelling highlight that 
a variety of actors have a direct interest in gaining such knowledge 
including network equipment suppliers; network operators, regulatory 
and local authorities. 
These processes and their effects very a lot across scales: although
the diffusion of a new technology might not be optimal at a local level,
it might be beneficial from a global perspective as it could lead to 
faster diffusion to less advantaged places [@leibowicz2016representing]. 
Despite the spatial heterogeneity of such diffusion mechanisms and 
the policy relevance, there are very limited attempts in the literature 
to analyse the diffusion of new digital technologies at a detailed 
geographical level.

@@

<!-- previous studies -->
The remaining of this section reviews empirical studies which analysed the diffusion 
on new technologies over space and time. Although the spatial dimension is present
in most of the following studies, the level of spatial detail is almost always 
more coarse than the one adopted here. To begin with @beardsell1999spatial studied 
the evolution of the computer industry in $317$ US metro areas during the $1977-1992$ 
period using employment data. Their analysis indicated that the relative size distribution
holds for urban computer employment and also urban heterogeneity is essential in
explaining this distribution. in a recent study, @bednarz2020pulled focused on 
wind turbines and modelled their spatial diffusion across $402$ German regions 
during $1970-2015$. Their key finding is that local demand than local supply was 
the main driving factor.

>>>>

<!-- motivation -->
The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffused over space and time and use this knowledge to make predictions
regarding the diffusion of related *future* technologies.
As per @leibowicz2016representing, historical studies agree on the fact that 
"technologies diffuse at different times, at different rates, and to
different extents in different places, and can be significantly
influenced by policies" [@victor1993]. @meade2021modelling highlight that 
a variety of actors have a direct interest in gaining such knowledge 
including network equipment suppliers; network operators, regulatory 
and local authorities. 
These processes and their effects vary a lot across scales: although
the diffusion of a new technology might not be optimal at a local level,
it might be beneficial from a global perspective as it could lead to 
faster diffusion to less advantaged places [@leibowicz2016representing]. 
Despite the spatial heterogeneity of such diffusion mechanisms and 
the policy relevance, there are very limited attempts in the literature 
to analyse the diffusion of new digital technologies at a detailed 
geographical level.

@@

<!-- previous studies -->
The remaining of this section reviews empirical studies which analysed the diffusion 
on new technologies over space and time. Although the spatial dimension is present
in most of the following studies, the level of spatial detail is almost always 
more coarse than the one adopted here. @beardsell1999spatial studied 
the evolution of the computer industry in $317$ US metro areas during the $1977-1992$ 
period using employment data. Their analysis indicated that the relative size distribution
holds for urban computer employment and also urban heterogeneity is essential in
explaining this distribution. in a recent study, @bednarz2020pulled focused on 
wind turbines and modelled their spatial diffusion across $402$ German regions 
during $1970-2015$. Their key finding is that local demand than local supply was 
the main driving factor.

### 2024-02-08 presentations/web_diffusion.Rmd
bua

    -   S-shaped diffusion curves
    -   Rank dynamics
    -   Spatial analysis
    -   Modelling

@@

-   Diffusion of an intangible/digital technology

@@

-   Over time, **early stages** of the internet

@@

<small>

-   Tranos, E., and C. Stich. 2020. Individual internet usage and the
    availability of online content of local interest: A multilevel
    approach. *Computers, Environment and Urban Systems*, 79:101371.

-   Tranos, E., T. Kitsos, and R. Ortega-Argilés. 2021. Digital economy
    in the UK: Regional productivity effects of early adoption.
    *Regional Studies*, 55:12, 1924-1938.

-   Stich, C., E. Tranos and M. Nathan. 2022. Modelling clusters from
    the ground up: a web data approach. *Environment and Planning B*, in
    press.

-   Tranos, E., A. C. Incera and G. Willis. 2022. Using the web to
    predict regional trade flows: data extraction, modelling, and
    validation, *Annals of the AAG*, in press.

</small>

@@

20080509162138/<http://www.website1.co.uk/contact_us> IG8 8HD

@@

-   Model cumulative adoption

@@

-   Modelling framework

@@

```{r LA_S_gif, eval=F, results='asis', out.width = "60%"}
path.images <- paste0(path, "/outputs/")

library(magick)
las <- list.files(path = path.images, pattern = "^lad_", full.names = T) #%>% 
las_list <- lapply(las, image_read)

@@

## animate at 2 frames per second
las_animated <- image_animate(las_joined, fps = 2)

## view animated image
las_animated

## save to disk
image_write(image = las_animated,
            path = "las.gif")

@@

@@

-   Website density regressions -- distance effect

@@

## Website density regressions

$$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$

::: {style="margin-left: 50px !important; text-align:left !important"}
$Website\,Density_{i}$:

-   Websites per firm in a Local Authority $i$, *or*

-   Websites in an Output Area $i$
:::

## Website density regressions

$$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$

::: {style="margin-left: 50px !important; text-align:left !important"}
$Place$:

-   London, *or*

-   Nearest city, *or*

-   Nearest retail centre
:::

## Website density regressions

$\beta$ interpretation:

-   The lower the $\beta$ is (or the larger the $|\beta|$ is)...

-   ... the larger urban gravitation is for web adoption.

```{r b_London_la, fig.height=5, fig.width=5, message=FALSE}
path.image <- paste0(path, "/outputs/lisa/b_dist_london_la.png")
knitr::include_graphics(path.image)
```

```{r b_city_la, fig.height=5, fig.width=5, message=FALSE}
#, out.width = "60%"
path.image <- paste0(path, "/outputs/lisa/b_dist_city_la.png")
knitr::include_graphics(path.image)
```

```{r b_centre_la,  fig.height=5, fig.width=5,message=FALSE}
#, out.height= "60%", out.width="60%"
path.image <- paste0(path, "/outputs/lisa/b_dist_centre_la.png")
knitr::include_graphics(path.image)
```

```{r b_London_oa, eval=T, fig.height=5, fig.width=5, message=FALSE}
#, out.width = "60%"
path.image <- paste0(path, "/outputs/lisa/b_dist_london_oa.png")
knitr::include_graphics(path.image)
```

```{r b_centre_oa, eval = T,  fig.height=5, fig.width=5,message=FALSE}
#fig.height=4, fig.width=4, 
#, out.height= "60%", out.width="60%"
path.image <- paste0(path, "/outputs/lisa/b_dist_city_oa.png")
knitr::include_graphics(path.image)
```

```{r b_city_oa, eval = T, fig.height=5, fig.width=5, message=FALSE}
#, out.width = "60%"
path.image <- paste0(path, "/outputs/lisa/b_dist_centre_oa.png")
knitr::include_graphics(path.image)
```

```{r r_squared, eval=T, results='asis'}
# This is for the RSquared of the different website density regressions.

path.in1 <- paste0(path, "/outputs/lisa/r2_all_oa.csv")
path.in2 <- paste0(path, "/outputs/lisa/r2_all_la.csv")

# read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
#   ggplot(aes(x=year)) +
#   geom_line(aes(y=London, color = "OA to London"), size=1) +
#   geom_line(aes(y=cities, color = "OA to nearest city"), size=1) +
#   geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) +
# 
#   geom_line(aes(y=London.LA, color = "LA to London"), size=1, linetype = 2) +
#   geom_line(aes(y=cities.LA, color = "LA to nearest city"), size=1, linetype = 2) +
#   geom_line(aes(y=retail.LA, color = "LA to nearest retail centre"), size=1, linetype = 2) +
# 
#   scale_color_discrete(name = "R-squared for regression:") +
#   #scale_linetype_manual(name = "Website density regressions111") 
#   ylab("R-squared") + 
#   ggtitle("Website density regressions") +
#   theme_minimal() +
#   theme(axis.title.x=element_blank(),
#         plot.title = element_text(hjust = 0.5))

read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
  pivot_longer(!year, names_to = "regression", values_to = "Rsquared") %>% 
  mutate(LA = if_else(str_ends(regression, ".LA"), " to LA", "to OA"),
         regression = str_remove_all(regression, ".LA"),
         line = paste0(regression, " ", LA)) %>% 
  ggplot() +
  geom_line(aes(x=year, y = Rsquared, color = line),
            size = 1) +
  scale_color_discrete(name = "R-squared for regression:") +
  #scale_linetype_manual(name = "Website density regressions111") 
  ylab("R-squared") + 
  ggtitle("Website density regressions") +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        plot.title = element_text(hjust = 0.5))

#   scale_color_manual(values = c("red", "blue", "green"))

```

@@

-   Spatial dependency
    -   Relatively small, consistent over time / scales
    -   London hot spot early on
    -   At local scale, consistent hotspots over time
    -   Granular analysis reveals other hotspots
-   Distance effect
    -   Urban gravitation increases over time and then drops
    -   Granular analysis: gravitation
    -   Lost explanatory power over time

@@

@@

-   4 sets of models:

    -   All the data ⟹ *variable importance*
    -   Train on one region, test on the rest ⟹ *spatial differences and
        similarities* of diffusion mechanisms
    -   For Local Authorities and Output Areas

@@

## Models trained on all data
<br>

|                   | RMSE  | RSquared | MAE   |
|-------------------|-------|----------|-------|
| Local Authorities | 0.022 | 0.908    | 0.012 |
| Output Areas      | 3.613 | 0.627    | 0.533 |

@@

path.image <- paste0(path, "/outputs/rf/varimp_LA.png")
knitr::include_graphics(path.image)
```

@@

path.image <- paste0(path, "/outputs/rf/varimp_OA.png")
knitr::include_graphics(path.image)
```

@@

```{r similarity_LA, eval = T, message=FALSE}
#fig.height=5, fig.width=5 
#, out.height= "60%", out.width="60%"
path.image <- paste0(path, "/outputs/rf/heatmap_LA.png")
knitr::include_graphics(path.image)
```

@@

```{r similarity_OA, eval = T, out.height= "70%", out.width="70%", message=FALSE}

path.image <- paste0(path, "/outputs/rf/heatmap_OA.png")
knitr::include_graphics(path.image)
```

@@

-   Some indications of a hierarchical diffusion

@@

>>>>

    -   S-shaped diffusion curves
    -   Rank dynamics
    -   Exploratory spatial analysis
    -   Modelling

@@

-   Diffusion of an **intangible**, digital technology [*web*]

@@

-   Over time, **early stages** of the internet [*1996-2012*]

@@

<!-- <small> -->

<!-- -   Tranos, E., and C. Stich. 2020. Individual internet usage and the -->

<!--     availability of online content of local interest: A multilevel -->

<!--     approach. *Computers, Environment and Urban Systems*, 79:101371. -->

<!-- -   Tranos, E., T. Kitsos, and R. Ortega-Argilés. 2021. Digital economy -->

<!--     in the UK: Regional productivity effects of early adoption. -->

<!--     *Regional Studies*, 55:12, 1924-1938. -->

<!-- -   Stich, C., E. Tranos and M. Nathan. 2022. Modelling clusters from -->

<!--     the ground up: a web data approach. *Environment and Planning B*, in -->

<!--     press. -->

<!-- -   Tranos, E., A. C. Incera and G. Willis. 2022. Using the web to -->

<!--     predict regional trade flows: data extraction, modelling, and -->

<!--     validation, *Annals of the AAG*, in press. -->

<!-- </small> -->

@@

20080509162138 \| <http://www.website1.co.uk/contact_us> \| IG8 8HD

@@

-   Cumulative adoption: Self-starting logistic growth model [`nls` and
    `SSlogis`]

@@

-   Machine learning framework [random forests]

@@

```{r LA_S_gif, eval=F, results='asis', out.width = "60%"}
path.images <- paste0(path, "/outputs/s_lad_per_firm/")

library(magick)
las <- list.files(path = path.images, pattern = "^lad_", full.names = T) 
las <- sample(las, 50)
las_list <- lapply(las, image_read)

@@

## animate at 2 frames per second
las_animated <- image_animate(las_joined, fps = 4)

## view animated image
#las_animated

## save to disk
# image_write(image = las_animated,
#             path = "las.gif")
image_write_gif(las_animated, path = "las.gif")

@@

```{r gif S, eval=T, out.width = "60%", out.height="60%"}
#, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"

path.image <- paste0(path, "/presentations/las.gif")
knitr::include_graphics(path.image)
```

@@

<!-- -   Website density regressions -- distance effect -->

@@

<!-- ## Website density regressions -->

<!-- <br> -->

<!-- $$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$ -->

<!-- <br> -->

<!-- ::: {style="margin-left: 50px !important; text-align:left !important"} -->

<!-- $Website\,Density_{i}$: -->

<!-- -   Websites per firm in a Local Authority $i$, *or* -->

<!-- -   Websites in an Output Area $i$ -->

<!-- ::: -->

<!-- ## Website density regressions -->

<!-- <br> -->

<!-- $$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$ -->

<!-- <br> -->

<!-- ::: {style="margin-left: 50px !important; text-align:left !important"} -->

<!-- $Place$: -->

<!-- -   London, *or* -->

<!-- -   Nearest city, *or* -->

<!-- -   Nearest retail centre -->

<!-- ::: -->

<!-- ## Website density regressions -->

<!-- <br> -->

<!-- $\beta$ interpretation: -->

<!-- -   The lower the $\beta$ is (or the larger the $|\beta|$ is)... -->

<!-- -   ... the larger urban gravitation is for web adoption. -->

<!-- ```{r b_London_la, fig.height=5, fig.width=5, message=FALSE} -->

<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_london_la.png") -->

<!-- knitr::include_graphics(path.image) -->

<!-- ``` -->

<!-- ```{r b_city_la, fig.height=5, fig.width=5, message=FALSE} -->

<!-- #, out.width = "60%" -->

<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_city_la.png") -->

<!-- knitr::include_graphics(path.image) -->

<!-- ``` -->

<!-- ```{r b_centre_la,  fig.height=5, fig.width=5,message=FALSE} -->

<!-- #, out.height= "60%", out.width="60%" -->

<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_centre_la.png") -->

<!-- knitr::include_graphics(path.image) -->

<!-- ``` -->

<!-- ```{r b_London_oa, eval=T, fig.height=5, fig.width=5, message=FALSE} -->

<!-- #, out.width = "60%" -->

<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_london_oa.png") -->

<!-- knitr::include_graphics(path.image) -->

<!-- ``` -->

<!-- ```{r b_centre_oa, eval = T,  fig.height=5, fig.width=5,message=FALSE} -->

<!-- #fig.height=4, fig.width=4,  -->

<!-- #, out.height= "60%", out.width="60%" -->

<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_city_oa.png") -->

<!-- knitr::include_graphics(path.image) -->

<!-- ``` -->

<!-- ```{r b_city_oa, eval = T, fig.height=5, fig.width=5, message=FALSE} -->

<!-- #, out.width = "60%" -->

<!-- path.image <- paste0(path, "/outputs/lisa/b_dist_centre_oa.png") -->

<!-- knitr::include_graphics(path.image) -->

<!-- ``` -->

<!-- ```{r r_squared, eval=T, results='asis'} -->

<!-- # This is for the RSquared of the different website density regressions. -->

<!-- path.in1 <- paste0(path, "/outputs/lisa/r2_all_oa.csv") -->

<!-- path.in2 <- paste0(path, "/outputs/lisa/r2_all_la.csv") -->

<!-- # read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>%  -->

<!-- #   ggplot(aes(x=year)) + -->

<!-- #   geom_line(aes(y=London, color = "OA to London"), size=1) + -->

<!-- #   geom_line(aes(y=cities, color = "OA to nearest city"), size=1) + -->

<!-- #   geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) + -->

<!-- #  -->

<!-- #   geom_line(aes(y=London.LA, color = "LA to London"), size=1, linetype = 2) + -->

<!-- #   geom_line(aes(y=cities.LA, color = "LA to nearest city"), size=1, linetype = 2) + -->

<!-- #   geom_line(aes(y=retail.LA, color = "LA to nearest retail centre"), size=1, linetype = 2) + -->

<!-- #  -->

<!-- #   scale_color_discrete(name = "R-squared for regression:") + -->

<!-- #   #scale_linetype_manual(name = "Website density regressions111")  -->

<!-- #   ylab("R-squared") +  -->

<!-- #   ggtitle("Website density regressions") + -->

<!-- #   theme_minimal() + -->

<!-- #   theme(axis.title.x=element_blank(), -->

<!-- #         plot.title = element_text(hjust = 0.5)) -->

<!-- read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>%  -->

<!--   pivot_longer(!year, names_to = "regression", values_to = "Rsquared") %>%  -->

<!--   mutate(LA = if_else(str_ends(regression, ".LA"), " to LA", "to OA"), -->

<!--          regression = str_remove_all(regression, ".LA"), -->

<!--          line = paste0(regression, " ", LA)) %>%  -->

<!--   ggplot() + -->

<!--   geom_line(aes(x=year, y = Rsquared, color = line), -->

<!--             size = 1) + -->

<!--   scale_color_discrete(name = "R-squared for regression:") + -->

<!--   #scale_linetype_manual(name = "Website density regressions111")  -->

<!--   ylab("R-squared") +  -->

<!--   ggtitle("Website density regressions") + -->

<!--   theme_minimal() + -->

<!--   theme(axis.title.x=element_blank(), -->

<!--         plot.title = element_text(hjust = 0.5)) -->

<!-- #   scale_color_manual(values = c("red", "blue", "green")) -->

<!-- ``` -->

@@

-   Spatial dependency
    -   Relatively small, consistent over time / scales
    -   London hot spot early on
    -   At local scale, consistent hotspots over time
    -   Granular analysis reveals other hotspots
        <!-- -   Distance effect -->
        <!--     -   Urban gravitation increases over time and then drops -->
        <!--     -   Granular analysis: gravitation -->
        <!--     -   Lost explanatory power over time -->

@@

-   [A **hierarchy effect**: from main centres to secondary ones]{style="color:orange;"}

-   [A **neighborhood effect**: first "hitting" nearby locations]{style="color:olive;"}

-   [**S-shaped** pattern in the cumulative level of adoption]{style="color:violet;"}

<br> <small>
$$Website\,Density_{t} \sim \color{orange}{Distance\,London} + 
\color{orange}{Website\,density\,London_{t-1}} + \\
\color{orange}{Distance\,Nearest\,City} + 
\color{orange}{Website\,density\,Nearest\,City_{t-1}} + \\
\color{orange}{Distance\,Nearest\,Retail_{i}} +
\color{orange}{Website\,density\,Nearest\,Retail_{t-1}} +\\
\color{olive}{W*\, Website\,density_{t-1}} +\\ 
\color{violet}{year_{t}}$$ </small>

@@

-   2 spatial resolutions:

    -   Local Authorities (websites per firm)
    -   Output Areas (websites)

-   2 sets of models:

    -   All the data ⟹ *variable importance*
    -   Train on all but one region, test on the holdout region ⟹
        *spatial differences and similarities* of diffusion mechanisms

@@

## Models trained on all data

|                   | RMSE  | RSquared | MAE   |
|-------------------|-------|----------|-------|
| Local Authorities | 0.032 | 0.810    | 0.019 |
| Output Areas*      | 5.000 | 0.205    | 1.047 |

<small>
*No retail centre variables for the Local Authority models
</small>

@@

path.image <- paste0(path, "/outputs/rf/figures/varimp_LA.png")
knitr::include_graphics(path.image)
```

@@

path.image <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
knitr::include_graphics(path.image)
```

@@

```{r similarity_LA, eval = T, message=FALSE}
path.table <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
read.csv(path.table) %>% 
  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
  kable()
```

@@

```{r similarity_OA, eval = F, out.height= "70%", out.width="70%", message=FALSE}

path.image <- paste0(path, "/outputs/rf/figures/heatmap_OA.png")
knitr::include_graphics(path.image)
```

## Alternative specification: growth, trained on all data

<br> <small>
$$Website\,Density\,Growth_{t} \sim \color{orange}{Distance\,London} + 
\color{orange}{Website\,density\,London_{t-1}} + \\
\color{orange}{Distance\,Nearest\,City} + 
\color{orange}{Website\,density\,Nearest\,City_{t-1}} + \\
\color{orange}{Distance\,Nearest\,Retail_{i}} +
\color{orange}{Website\,density\,Nearest\,Retail_{t-1}} +\\
\color{olive}{W*\, Website\,density_{t-1}} +\\ 
\color{violet}{year_{t}}$$ </small>

|                   | RMSE  | RSquared | MAE   |
|-------------------|-------|----------|-------|
| Local Authorities* | 0.200 | 0.634    | 0.148 |

<small>
*No retail centre variables for the Local Authority models
</small>

## Alternative specification: growth, trained on all data

```{r growth LA}
path.table <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth.csv")
read.csv(path.table) %>% 
  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
  kable()
```

@@

-   Hierarchical diffusion

@@

-   Things to do: OA?
